---
title: "Quandl"
author: "Data Collection for Biz Analytics @ ISB"
date: "February 20, 2018"
output: html_document
---

Quandl provides free and paid access to various quality financial data across the globe through its platform. We can get data from Quandl with API access or with R or with Python or simply excel. To know more about capabilities and usage please refer to [API documentation](https://docs.quandl.com/docs).

Quandl can provide data for up to 50 requests a day without using an API key. After that it will ask for API key. You can generate your free API key very easily. Just sign up for a free [Quandl](https://www.quandl.com/) account and login to your account. Now go to [account settings page]( https://www.quandl.com/account/api). In this page you can find the free API key. Copy this key and we will this API key later in R session.

Now next step is find the data source from which we want to download the data. You can see the full list of available data sources [here](https://www.quandl.com/search?query=). If desired data source is a premium data source then we have to subscribe for that data source.

Let's say we want to download stock prices data from National Stock Exchange, India. The corresponding data source name in Quandl is [NSE](https://www.quandl.com/data/NSE-National-Stock-Exchange-of-India). In this page we can click on usage tab and see all the available query methods for this data sources. For example see [R usage method](https://www.quandl.com/data/NSE-National-Stock-Exchange-of-India/usage/quickstart/r)

To get the data in R, we will use *Quandl* package in R. Next thing we need is to setup our API key in R session. We can setup API key in R session with *Quandl.api_key* function. See code below - 
```{r , message= F}
#install.packages("Quandl")
library(Quandl)
Quandl.api_key("6X-TZmhzrTjgJH29YDrk")
```

Now let's say we want to download the daily stock prices for TATA Motors (NSE ticker - TATAMOTORS). As shown in the R usage guide, we can use *Quandl* function to get this data. Please see the code below -

```{r }
tata_motors = Quandl('NSE/TATAMOTORS', start_date='2018-01-01', end_date='2018-02-19')
head(tata_motors)
```


```{r }
```
