---
title: "Data Collection from News API"
output: html_document
---
* Team Members :
    - Anmol More - 11915043
    - Dharani Kiran - 11915033
    - Shubhendu Sharma - 11915067

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


**Step 1 - Signup for an API key at https://newsapi.org/account.**  

Refer to accessible endpoints - https://newsapi.org/docs/endpoints

**Step 2 - Load required packages.**   

We will need `httr`, `RCurl` and `jsonlite` packages in the R Workspace. 

```{r import libraries, warning=FALSE}
suppressPackageStartupMessages({
  if (!require(httr)) {install.packages("httr")}; library(httr)
  if (!require(RCurl)) {install.packages("RCurl")}; library(RCurl)
  if (!require(jsonlite)) {install.packages("jsonlite")}; library(jsonlite)
})

```

**Step 3 - As mentioned in doc https://newsapi.org/docs/endpoints/everything**
** Hit endpoint with search parameter and api key**

```{r querying business analytics news}

search.term <- "business%20analytics"
api.key <- "8d88c21f6fd7423b9c84c38f78d84703"
page.size <- 100
url <- paste0("https://newsapi.org/v2/everything?q=", search.term, "&apiKey=", api.key, "&pageSize=", page.size)
print(url)

system.time({
  all.news <- getURL(url) 
})

news.json <- jsonlite::fromJSON(all.news)

```

**Iterate through news results and create a dataframe**

```{r create dataframe}

source.id <- news.json$articles$source$id
source.name <- news.json$articles$source$name
author <- news.json$articles$author
title <- news.json$articles$title
description <- news.json$articles$description
content <- news.json$articles$content

news.df <- data.frame(source.id, source.name, author, title, description, content)
headers <- c("Source Id", "Source Name", "Author", "Title", "Description", "Content")
colnames(news.df) <- headers

head(news.df)

```

**Write fetched news results to output file**

```{r generate csv}
write.csv(news.df, "output_api.csv")
```

